This page and the next page should be removed

### 2021-04-14 analysis/manuscript.Rmd
make dockerfile ready for binder
start with merge exps

@@

```{r setup, include=FALSE}
# Do you want to re-run the EEG preprocessing? Note that this can only be done with the raw data,
# which are available from the corresponding author (rasha.abdel.rahman@hu-berlin.de) upon request.
rerun_preproc <- FALSE

@@

# Define ERP components of interest
comps <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(100, 150, 400),
  tmax = c(150, 200, 700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)
```

@@

```{python def_functions, eval=rerun_preproc}
# Import libraries
import mne

# Define function for preprocessing the EEG data
def preproc(vhdr_fname, metadata, n_components, method, l_freq, h_freq, event_id, tmin, tmax,
    baseline, reject):
    # Read EEG data in microvolts
    raw = mne.io.read_raw_brainvision(vhdr_fname, scale=1e6, preload=True)
    # Create virtual EOG channels
    raw = mne.set_bipolar_reference(raw, "Auge_u", "Fp1", ch_name="VEOG", drop_refs=False)
    raw = mne.set_bipolar_reference(raw, "F9", "F10", ch_name="HEOG", drop_refs=False)
    raw.set_channel_types(mapping={"VEOG": "eog", "HEOG": "eog"})
    # Add EasyCap electrode layout, removing any excessive channels
    montage = mne.channels.make_standard_montage("easycap-M1")
    raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(["VEOG", "HEOG"])))
    raw.set_montage(montage=montage)
    # Re-reference to common average
    raw, _ = mne.set_eeg_reference(raw, "average")
    # Run ICA on a copy of the data
    filt_raw = raw.copy()
    filt_raw.load_data().filter(l_freq=1, h_freq=None)
    ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
    ica.fit(filt_raw)
    # Remove bad components based on correlations with EOG
    eog_indices, eog_scores = ica.find_bads_eog(raw)
    ica.exclude = eog_indices
    raw = ica.apply(raw)
    # Apply band-pass filter
    raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
    # Epoching including baseline correction
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(
        raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True
    )
    # Add behavioral data
    epochs.metadata = metadata
    # Reject bad epochs
    epochs = epochs.drop_bad(reject=reject)
    return epochs

# Define function to compute averaged waveforms by part and condition
def compute_evokeds(epochs, grand=True, data_only=True):
    # Define lists of participants, parts, and conditions
    ps = epochs.metadata["participant"].unique()
    pts = ["I", "II", "III"]
    cns = ["Informed", "Naive"]
    # Define function to compute a single by-participant waveform
    def compute_evoked(epochs, pt, cn, p):
        query = 'participant == "' + p + '" and part == "' + pt + '" and condition == "' + cn + '"'
        return epochs[query].average()
    # Create a dict with by-participant waveforms nested in conditions nested in parts
    avgs = {pt: {cn: [compute_evoked(epochs, pt, cn, p) for p in ps] for cn in cns} for pt in pts}
    # If needed, compute grand-averaged waveforms across participants
    if grand:
        avgs = {pt: {cn: mne.grand_average(avgs[pt][cn]) for cn in cns} for pt in pts}
    # If needed, extract only the data from the evoked MNE objects
    if data_only:
        if grand:
            avgs = {pt: {cn: avgs[pt][cn].data for cn in cns} for pt in pts}
        else:
            ps_num = range(0, len(ps))
            avgs = {pt: {cn: [avgs[pt][cn][p].data for p in ps_num] for cn in cns} for pt in pts}
    return avgs

# Define function to compute single-trial mean ERP amplitudes from epochs
def compute_erps(epochs, name, tmin, tmax, roi):
    # Average across electrodes in the ROI
    roi_idx = mne.pick_channels(epochs.ch_names, roi)
    epochs_roi = mne.channels.combine_channels(epochs, dict(roi=roi_idx))
    # Average across samples in the time window
    import numpy as np
    tmin_s, tmax_s = tmin / 1000, tmax / 1000
    times_idx = np.where(np.logical_and(epochs.times >= tmin_s, epochs.times <= tmax_s))
    erps = epochs_roi.get_data()[:, 0, times_idx].mean(axis=(1, 2))
    return erps

@@

# Preprocess the EEG for Experiment 1 in Python
map2(
  .x = list.files("data/exp1/EEG", pattern = ".vhdr", full.names = TRUE),
  .y = behav_exp1,
  ~ do.call(py$preproc, c(list(vhdr_fname = .x, metadata = .y), preproc_params))
) %>%
  # Concatenate epochs of all participants
  py$mne$epochs$concatenate_epochs() -> epochs_exp1

# Backup epochs to the export folder
epochs_exp1$save("analysis/export/exp1-epo.fif")

# Compute mean ERP amplitudes and bind to the behavioral data
pmap_dfc(comps, py$compute_erps, epochs = epochs_exp1) %>%
  set_names(comps$name) %>%
  bind_cols(as_tibble(epochs_exp1$metadata), .) %T>%
  write_tsv("analysis/export/exp1-data.tsv") -> data_exp1

# Compute grand averages per part and condition
py$compute_evokeds(epochs = epochs_exp1, grand = TRUE, data_only = TRUE) %>%
  saveRDS("analysis/export/exp1-gaverages.RDS")

# Backup number of stimuli assigned to each condition for all participants
map(behav_exp1, function(x) table(x$condition) / 3) %>%
  bind_rows() %>%
  mutate(Participant = unique(data_exp1$participant), .before = 1) %>%
  write_tsv("analysis/export/exp1-objects.tsv")

# Get standard EasyCap electrode positions and project to 2D space
py$mne$channels$`_standard_montage_utils`$MONTAGE_PATH %>%
  list.files(pattern = "easycap-M1.txt", full.names = TRUE) %>%
  read_tsv(col_types = cols()) %>%
  rename(electrode = Site) %>%
  inner_join(tibble(electrode = epochs_exp1$ch_names), ., by = "electrode") %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)),
  ) %>%
  write_tsv("analysis/export/montage.tsv")

# Remove the epochs to free up memory
rm(epochs_exp1)
```

```{r exp1-models, eval=rerun_models}
# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known"
    )),
    participant = factor(participant),
    item = factor(item)
  ) %>%
  # Remove any trials excluded from conditions
  filter(condition %in% c("Informed", "Naive")) %>%
  droplevels() -> data_exp1

@@

# Apply these contrasts to the factors in the data
contrasts(data_exp1$condition) <- MASS::ginv(contrasts_condition)
contrasts(data_exp1$part) <- MASS::ginv(contrasts_part)

@@

# Compute the models for Experiment 1
map(
  comps$name,
  compute_models,
  formula = formula_exp12,
  data = data_exp1,
  control = control,
  specs = specs_exp12
) %>%
  set_names(comps$name) %T>%
  saveRDS("analysis/export/exp1-models.RDS") -> models_exp1
```

@@

# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")
    ),
    participant = factor(participant),
    item = factor(item)
  ) -> data_exp1

# Check number of rejected epochs
data_exp1 %>%
  group_by(participant) %>%
  tally() %>%
  mutate(n = 360 - n) %>%
  pull(n) -> rejected_exp1
mean(rejected_exp1)    # Mean
median(rejected_exp1)  # Median
range(rejected_exp1)   # Range

@@

# epochs = preproc(
#   vhdr_fname="data/exp1/eeg/Vp0004.vhdr",
#   metadata=r.data_exp1.query("participant == 'VP_04'"),
#   n_components=15,
#   method="fastica",
#   l_freq=None,
#   h_freq=None,
#   event_id=dict(match=221, mismatch=222),
#   tmin=-0.5,
#   tmax=1.498,
#   baseline=(-0.2, 0),
#   reject=dict(eeg=200)
# )
# epochs.save("analysis/export/vp04-epo.fif")

@@

```{r joint-import}
# Combine data from Experiments 1 and 2 (requires changing the participant IDs for Experiment 2)
data_exp2 %>%
  mutate(participant = fct_relabel(participant, ~ paste0(., "_2"))) %>%
  bind_rows(data_exp1, ., .id = "experiment") %>%
  mutate(experiment = as_factor(experiment)) -> data_joint

>>>>

editor_options: 
  chunk_output_type: console

@@

```{r setup, include=FALSE}
# Do you want to re-run the EEG preprocessing? Note that this can only be done with the raw data,
# which are available from the corresponding author (rasha.abdel.rahman@hu-berlin.de) upon request.
rerun_preproc <- TRUE

@@

# Define ERP components of interest
comps <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(100, 150, 400),
  tmax = c(150, 200, 700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

# Define some filenames for I/O
fnames <- list(
  epochs = "analysis/export/aha-epo.fif",
  trials = "analysis/export/aha-trials.csv",
  gavgs = "analysis/export/aha-gavgs.RDS",
  conditions = "analysis/export/aha-obj-per-cond.csv",
  montage = "analysis/export/aha-montage.csv"
)
```

@@

```{python def_functions, eval=rerun_preproc}
# Import libraries
import mne
import numpy as np

# Define function for preprocessing a single participant
def preproc(
    vhdr_fname,
    metadata,
    n_components,
    method,
    l_freq,
    h_freq,
    event_id,
    tmin,
    tmax,
    baseline,
    reject,
):
    # Read EEG data in microvolts
    raw = mne.io.read_raw_brainvision(vhdr_fname, scale=1e6, preload=True)
    # Create virtual EOG channels
    raw = mne.set_bipolar_reference(raw, "Auge_u", "Fp1", ch_name="VEOG", drop_refs=False)
    raw = mne.set_bipolar_reference(raw, "F9", "F10", ch_name="HEOG", drop_refs=False)
    raw.set_channel_types(mapping={"VEOG": "eog", "HEOG": "eog"})
    # Add EasyCap electrode layout, removing any excessive channels
    montage = mne.channels.make_standard_montage("easycap-M1")
    raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(["VEOG", "HEOG"])))
    raw.set_montage(montage=montage)
    # Re-reference to common average
    raw, _ = mne.set_eeg_reference(raw, "average")
    # Run ICA on a copy of the data
    filt_raw = raw.copy()
    filt_raw.load_data().filter(l_freq=1, h_freq=None)
    ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
    ica.fit(filt_raw)
    # Remove bad components based on correlations with EOG
    eog_indices, eog_scores = ica.find_bads_eog(raw)
    ica.exclude = eog_indices
    raw = ica.apply(raw)
    # Apply band-pass filter
    raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
    # Epoching including baseline correction
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(
        raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True
    )
    # Add behavioral data
    epochs.metadata = metadata
    # Reject bad epochs
    epochs = epochs.drop_bad(reject=reject)
    return epochs

# Define function for preprocessing a list of participants
def preproc_all(
    vhdr_fnames,
    metadatas,
    output_fname,
    n_components,
    method,
    l_freq,
    h_freq,
    event_id,
    tmin,
    tmax,
    baseline,
    reject,
):
    # Do the actual preprocessing
    epochs = [
        preproc(
            vhdr_fname=vhdr_fname,
            metadata=metadata,
            n_components=n_components,
            method=method,
            l_freq=l_freq,
            h_freq=h_freq,
            event_id=event_id,
            tmin=tmin,
            tmax=tmax,
            baseline=baseline,
            reject=reject,
        )
        for vhdr_fname, metadata in zip(vhdr_fnames, metadatas)
    ]
    # Concatenate all participants and save
    epochs = mne.epochs.concatenate_epochs(epochs)
    epochs.save(output_fname, overwrite=True)
    # Return the epochs without loading the actual data
    epochs = mne.read_epochs(output_fname, preload=False)
    return epochs

# Define function to compute averaged waveforms by part and condition
def compute_evokeds(epochs_fname, grand=True, data_only=True):
    # Read preprocessed epochs
    epochs = mne.read_epochs(epochs_fname)
    # Define lists of participants, parts, and conditions
    ps = epochs.metadata["participant"].unique()
    pts = ["I", "II", "III"]
    cns = ["Informed", "Naive"]
    # Define function to compute a single by-participant waveform
    def compute_evoked(epochs, pt, cn, p):
        query = 'participant == "' + p + '" and part == "' + pt + '" and condition == "' + cn + '"'
        return epochs[query].average()

    # Create a dict with by-participant waveforms nested in conditions nested in parts
    avgs = {pt: {cn: [compute_evoked(epochs, pt, cn, p) for p in ps] for cn in cns} for pt in pts}
    # If needed, compute grand-averaged waveforms across participants
    if grand:
        avgs = {pt: {cn: mne.grand_average(avgs[pt][cn]) for cn in cns} for pt in pts}
    # If needed, extract only the data from the evoked MNE objects
    if data_only:
        if grand:
            avgs = {pt: {cn: avgs[pt][cn].data for cn in cns} for pt in pts}
        else:
            ps_num = range(0, len(ps))
            avgs = {pt: {cn: [avgs[pt][cn][p].data for p in ps_num] for cn in cns} for pt in pts}
    return avgs

# Define function to compute single-trial mean ERP amplitudes from epochs
def compute_erps(epochs_fname, name, tmin, tmax, roi):
    # Read preprocessed epochs
    epochs = mne.read_epochs(epochs_fname)
    # Average across electrodes in the ROI
    roi_idx = mne.pick_channels(epochs.ch_names, roi)
    epochs_roi = mne.channels.combine_channels(epochs, dict(roi=roi_idx))
    # Average across samples in the time window
    tmin_s, tmax_s = tmin / 1000, tmax / 1000
    times_idx = np.where(np.logical_and(epochs.times >= tmin_s, epochs.times <= tmax_s))
    erps = epochs_roi.get_data()[:, 0, times_idx].mean(axis=(1, 2))
    return erps

@@

# Preprocess the EEG for Experiment 1 in Python
paste0("data/eeg/exp", 1:2) %>%
  list.files(pattern = ".vhdr", full.names = TRUE) %>%
  exec(
    .fn = py$preproc_all,
    vhdr_fnames = .,
    metadatas = behav,
    output_fname = fnames$epochs,
    !!!preproc_params
  ) -> epochs

# Compute mean ERP amplitudes and bind to the behavioral data
pmap_dfc(comps, py$compute_erps, epochs_fname = fnames$epochs) %>%
  set_names(comps$name) %>%
  bind_cols(epochs$metadata, .) %T>%
  write_tsv(fnames$trials) -> aha

# Compute grand averages per part and condition
py$compute_evokeds(epochs_fname = fnames$epochs, grand = TRUE, data_only = TRUE) %>%
  saveRDS(fnames$gavgs)

# Backup number of stimuli assigned to each condition for all participants
map(behav, function(x) table(x$condition) / 3) %>%
  bind_rows() %>%
  mutate(Participant = unique(aha$participant), .before = 1) %>%
  write_csv(fnames$conditions)

# Get standard EasyCap electrode positions and project to 2D space
py$mne$channels$`_standard_montage_utils`$MONTAGE_PATH %>%
  list.files(pattern = "easycap-M1.txt", full.names = TRUE) %>%
  read_tsv(col_types = cols()) %>%
  rename(electrode = Site) %>%
  inner_join(tibble(electrode = epochs$ch_names), ., by = "electrode") %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)),
  ) %>%
  write_csv(fnames$montage)
```

```{r exp1-models, eval=rerun_models}
# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known"
    )),
    participant = factor(participant),
    item = factor(item)
  ) %>%
  # Remove any trials excluded from conditions
  filter(condition %in% c("Informed", "Naive")) %>%
  droplevels() -> aha

@@

# Apply these contrasts to the factors in the data
contrasts(aha$condition) <- MASS::ginv(contrasts_condition)
contrasts(aha$part) <- MASS::ginv(contrasts_part)

@@

# Compute the models for Experiment 1
map(
  comps$name,
  compute_models,
  formula = formula_exp12,
  data = aha,
  control = control,
  specs = specs_exp12
) %>%
  set_names(comps$name) %T>%
  saveRDS("analysis/export/exp1-models.RDS") -> models_exp1
```

@@

# Load preprocessed single-trial data
read_tsv("analysis/export/exp1-data.tsv", col_types = cols()) %>%
  # Factorize columns
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    condition = factor(condition, levels = c(
      "Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")
    ),
    participant = factor(participant),
    item = factor(item)
  ) -> aha

# Check number of rejected epochs
aha %>%
  group_by(participant) %>%
  tally() %>%
  mutate(n = 360 - n) %>%
  pull(n) -> rejected_exp1
mean(rejected_exp1)    # Mean
median(rejected_exp1)  # Median
range(rejected_exp1)   # Range

@@

# epochs = preproc(
#   vhdr_fname="data/exp1/eeg/Vp0004.vhdr",
#   metadata=r.aha.query("participant == 'VP_04'"),
#   n_components=15,
#   method="fastica",
#   l_freq=None,
#   h_freq=None,
#   event_id=dict(match=221, mismatch=222),
#   tmin=-0.5,
#   tmax=1.498,
#   baseline=(-0.2, 0),
#   reject=dict(eeg=200)
# )
# epochs.save("analysis/export/vp04-epo.fif")

@@

```{r joint-import}
# Combine data from Experiments 1 and 2 (requires changing the participant IDs for Experiment 2)
data_exp2 %>%
  mutate(participant = fct_relabel(participant, ~ paste0(., "_2"))) %>%
  bind_rows(aha, ., .id = "experiment") %>%
  mutate(experiment = as_factor(experiment)) -> data_joint
